### 2022-03-22 misc/expo/fourier.tex
Couple of mistakes

\proof Write $Z \cong \ZZ_{n_1} \times \ZZ_{n_2}\times \cdots\times \ZZ_{n_r}$. For every
$u = (u_1,\ldots, u_r)\in Z$, the function $\chi_u : Z\to \CC$ that takes
$$(x_1,\ldots, x_r) \mapsto \prod_{i=1}^r e\biggl( {u_ix_i\over n_i}\biggr)$$
is a character (this is easy to show, since $e(x_i + x_i') = e(x_i)e(x_i')$). If $u\ne u'$,
then $\chi_u\ne\chi_{u'}$ and $\langle \chi_u,\chi_{u'}\rangle$ by the previous lemma. Hence
the set $\{\chi_u : u\in Z\}$ comprises $|Z|$ linearly independent elements in a space of dimension $|Z|$,
which must be spanning.\slug

@@

If $A$ is a subset of a finite abelian group $Z$, we associate to $A$ the {\it characteristic function}
$\one_A : Z\to \CC$ given by
$$\one_A(x) = \cases{ 1, & if $x\in A$;\cr 0, & otherwise.}$$
When it will not cause confusion, we will abuse notation and write $A(x)$ instead of $\one_A(x)$.
With the definitions and results stated above, we are now in a position to translate properties of $A$ to
Fourier-analytic statements about the function $\one_A$. First off, note that
$$\bignorm{\hat{\one_A}}_2^2 = \norm{\one_A}_2^2 = \ex_{x\in Z} \one_A(x)^2 = \ex_{x\in Z} \one_A(x)
= {|A|\over |Z|}.$$
If $|A|/|Z| = \delta$, then we will say that $A$ has {\it density} $\delta$. Another way of expressing $\delta$
in terms of $\one_A$ is
$$\hat{\one_A}(\chi_0) = \ex_{x\in Z} \one_A(x)\chi_0(x) = \ex_{x\in Z} \one_A(x) = \delta,$$
where $\chi_0$ is the trivial character.

The fact that $\langle \one_A, \one_A\rangle = \delta$ means that if we sample $x$ and $y$ from $Z$ and
condition that $x=y$, the probability that $x$ (or $y$) is in $A$ is $\delta$. The reason for stating this
obvious fact in such a stilted way is that it generalises by way of the convolution operation we defined
earlier; that is, what if we are interested in the probability that a $4$-tuple $(x,y,z,w)$ satisfying $x+y=z+w$
is a member of $A^4$? Well, the convolution law gives
$$\eqalign{
\ex_{x+y=z+w} A(x)A(y)A(z)A(w) &= \ex_{u\in Z} \ex_{x+y=u}\ex_{z+w=u} A(x)A(y)A(z)A(w) \cr
&= \ex_{u\in Z} (\one_A*\one_A)(u)^2 \cr
&= \norm{\one_A*\one_A}_2^2 \cr
&= \bignorm{\hat{\one_A*\one_A}}_2^2 \cr
&= \bignorm{\bigl(\hat{\one_A}\bigr)^2}_2^2 \cr
&= \sum_{\chi\in \hat Z} \bigl| \hat{\one_A}(\chi)\bigr|^4 \cr
&= \norm{\one_A}_4^4, \cr
}$$
and more generally, the probability that a tuple $(a_1,\ldots,a_k, b_1,\ldots,b_k)\in Z^{2k}$
with $a_1+\cdots+a_k=b_1+\cdots+b_k$ is a member of $A^{2k}$ is $\norm{\one_A}_{2k}^{2k}$. In any case,
the number of $(x,y,z,w)\in A^4$ with $x+y = z+w$ is so important in additive combinatorics that is has a name.
It is called the {\it additive energy} of $A$ and is denoted $E(A)$; what we showed above is that
$E(A) = |Z|^3 \norm{\one_A}_4^4$.

>>>>

\proof Write $Z \cong \ZZ_{n_1} \times \ZZ_{n_2}\times \cdots\times \ZZ_{n_r}$. For every
$u = (u_1,\ldots, u_r)\in Z$, the function $\chi_u : Z\to \CC$ that takes
$$(x_1,\ldots, x_r) \mapsto \prod_{i=1}^r e\biggl( {u_ix_i\over n_i}\biggr)$$
is a character (this is easy to show, since $e(x_i + x_i') = e(x_i)e(x_i')$). If $u\ne u'$,
then $\chi_u\ne\chi_{u'}$ and $\langle \chi_u,\chi_{u'}\rangle = 0$ by the previous lemma. Hence
the set $\{\chi_u : u\in Z\}$ comprises $|Z|$ linearly independent elements in a space of dimension $|Z|$,
which must be spanning.\slug

@@

If $A$ is a subset of a finite abelian group $Z$, we associate to $A$ the {\it characteristic function}
$\one_A : Z\to \CC$ given by
$$\one_A(x) = \cases{ 1, & if $x\in A$;\cr 0, & otherwise.}$$
When it will not cause confusion, we will abuse notation and write $A(x)$ instead of $\one_A(x)$.
With the definitions and results stated above, we are now in a position to translate properties of $A$ to
Fourier-analytic statements about the function $\one_A$. First off, note that
$$\bignorm{\hat{\one_A}}_2^2 = \norm{\one_A}_2^2 = \ex_{x\in Z} A(x)^2 = \ex_{x\in Z} A(x)
= {|A|\over |Z|}.$$
If $|A|/|Z| = \delta$, then we will say that $A$ has {\it density} $\delta$. Another way of expressing $\delta$
in terms of $\one_A$ is
$$\hat{\one_A}(\chi_0) = \ex_{x\in Z} A(x)\chi_0(x) = \ex_{x\in Z} A(x) = \delta,$$
where $\chi_0$ is the trivial character.

The fact that $\norm{\one_A}_2^2 = \delta$ means that if we sample $x$ and $y$ from $Z$ and
condition that $x=y$, the probability that $x$ (or $y$) is in $A$ is $\delta$. The reason for stating this
obvious fact in such a stilted fashion is that it generalises by way of the convolution operation we defined
earlier; that is, what if we are interested in the probability that a $4$-tuple $(x,y,z,w)$ satisfying $x+y=z+w$
is a member of $A^4$? Well, the convolution law gives
$$\eqalign{
\ex_{x+y=z+w} A(x)A(y)A(z)A(w) &= \ex_{u\in Z} \ex_{x+y=u}\ex_{z+w=u} A(x)A(y)A(z)A(w) \cr
&= \ex_{u\in Z} (\one_A*\one_A)(u)^2 \cr
&= \norm{\one_A*\one_A}_2^2 \cr
&= \bignorm{\hat{\one_A*\one_A}}_2^2 \cr
&= \bignorm{\bigl(\hat{\one_A}\bigr)^2}_2^2 \cr
&= \sum_{\chi\in \hat Z} \bigl| \hat{\one_A}(\chi)\bigr|^4 \cr
&= \norm{\one_A}_4^4, \cr
}$$
and more generally, the probability that a tuple $(a_1,\ldots,a_k, b_1,\ldots,b_k)\in Z^{2k}$
with $a_1+\cdots+a_k=b_1+\cdots+b_k$ is a member of $A^{2k}$ is $\norm{\one_A}_{2k}^{2k}$. In any case,
the number of $(x,y,z,w)\in A^4$ with $x+y = z+w$ is so important in additive combinatorics that is has a name.
It is called the {\it additive energy} of $A$ and is denoted $E(A)$; we just showed above is that
$E(A) = |Z|^3 \norm{\one_A}_4^4$.
